---
title: "IRD Analysis"
author: "Maliha"
date: "2023-09-12"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries}
library(dplyr)
library(tidyr)
library(ggplot2)
library(broom)
library(car)
library(gtsummary)
library(lubridate)
library(readxl)
library(stringr)


```

```{r}
screendata <- read.csv("Datasets/PSZ Impact Screening Data.csv", na.strings = "")
sessiondata <- read.csv("Datasets/PSZ Impact Session Data.csv", na.strings = "")

```
#Insert tables
```{r}

```

```{r}
# Change the format of the date variables to keep the earlier screening date
# Arranged the variables in desirable order
# Removed the duplicates in screendata

screendata <- screendata %>% 
  mutate(DATE_ENTERED_SCREENING = as_date(DATE_ENTERED_SCREENING, format = '%m/%d/%y')) %>% 
  arrange(PATIENT_ID, DATE_ENTERED_SCREENING) %>% 
  distinct(DATE_ENTERED, USER_ID, PATIENT_ID, loc_facility, GENDER, .keep_all = TRUE)

# Merged session data and screen data

df <- left_join(sessiondata, screendata) 

```


```{r}
# Rename variables and Reassign data frame with selected variables

df <- df %>% 
  select(ID = PATIENT_ID, AGE, GENDER, Marital_status = PART_MARI,
         EDUCATION, loss_of_pleasure, worry, difficulty_in_controlling_worry,
         feeling_depressed, PHQ4_score, FIRST_TOTAL_SCORE_PHQ, LAST_TOTAL_SCORE_PHQ,
         starts_with('DATE_FOLLOWUP_'), Status = TermSt)  

# Make variables all lower case for consistency
         
colnames(df) <- str_to_lower(colnames(df))

```

```{r}
df$status <- factor(df$status, 
                    levels = c('Completed therapy',
                               'Loss to Follow-up',
                               'Refused therapy',
                               'Referred',
                               'Other.'), 
                    ordered = TRUE)

df <- df %>% mutate(edu_cat = case_match(education, 
                        c('Primary (Class 1-5)', 'Secondary (Class 6-8)') ~ 'Primary',
                        'Matric (9-10)' ~ 'Matric',
                        'Intermediate (11-12)' ~ 'Intermediate',
                        'Bachelors' ~ 'Bachelors',
                        'Masters' ~ 'Masters', 
                        'Kuch nahi' ~ 'No Education',
                        'Islamic education' ~ 'Islamic Education'))

df$edu_cat <- factor(df$edu_cat,
                     levels = c('Islamic Education',
                                'No Education',
                                'Primary',
                                'Matric',
                                'Intermediate',
                                'Bachelors',
                                'Masters'),
                     ordered = TRUE)
                                
                        
table(df$edu_cat)       

```
```{r}
#Combine Marital Status Divorced and Separated 
table(df$marital_status)
df <- df %>% mutate(mari_cat = case_match(marital_status, 
                                          c('Divorced','Separated') ~ 'Divorced/Separated',
                                          'Married' ~ 'Married',
                                          'Single' ~ 'Single',
                                          'Widowed' ~ 'Widowed'))

df$mari_cat <- factor(df$mari_cat,
                     levels = c('Single',
                                'Married',
                                'Divorced/Separated',
                                'Widowed'),
                     ordered = TRUE)

table(df$mari_cat)

```


```{r}
##table1 <- CreateTableOne(data=df, factorVars = c('gender', 'marital_status', 'edu_cat'), vars = c('gender', 'marital_status', 'edu_cat', 'age'))
##kableone(table1)
#table1(~factor(gender) + age + factor(marital_status), data=df)

df |> select(age, gender, edu_cat, mari_cat, status) %>% 
  tbl_summary() %>% 
  as_gt()

df |> 
  select(age, gender, edu_cat, mari_cat, status) |> 
  tbl_summary(
    label = c(age = "Age",
              gender = "Gender",
              edu_cat = "Education",
              mari_cat = "Marital Status",
              status = "Therapy Status")) 

#how should I deal with the unknowns? leave them as it is?
# df |> 
#   select(age, gender, edu_cat, marital_status, diagnosis, status) |> 
#   tbl_summary(
#     by = diagnosis,
#     label = c(age = "Age",
#               gender = "Gender",
#               edu_cat = "Education",
#               marital_status = "Marital Status",
#               status = "Therapy Status")) |> 
#   add_overall(last = TRUE) |> 
#   add_p()




```



```{r}
table(df$status, useNA = 'always')
table(df$marital_status, useNA = 'always')
table(df$education, useNA = 'always')

df %>% ggplot() +
  geom_bar(aes(x = mari_cat))

statusplot <- df %>% ggplot() +
  geom_bar(aes(x = status))
statusplot
ggsave('Plots/statusplot.pdf', statusplot, height = 5, width = 7)

```
```{r}
df %>% ggplot() +
  geom_bar(aes(x = gender, fill = status))

df %>% ggplot(aes(x = status, group = gender)) +
  geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat = 'count') +
  facet_grid(~gender) +
  theme(legend.position = 'none') + 
  labs(y = 'Proportion', x = 'Therapy Status', title = 'Distribution of Gender by Status') +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  theme(plot.title = element_text(hjust = 0.5)) 

#ggsave('Plots/statusplot.pdf', statusplot, height = 5, width = 7)

df %>% ggplot() +
  geom_bar(aes(x = gender, fill = status), 
           position = 'fill') + 
  labs(y = 'Proportion', title = 'Distribution of Gender by Status') +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_brewer(palette = 'RdPu') 
```

```{r}
df <- df %>%
  mutate(diagnosis = case_when(phq4_score <= 6 ~ 'asymtomptic',
                               phq4_score > 6 ~ 'symtomptic'))

diagnosis_table <- table(df$gender, df$diagnosis)


chisq.test(diagnosis_table)

#linear regression
depression_model <- lm(phq4_score ~ gender, data = df)
summary(depression_model)

#the mean depression for female is ---- and male is ---, conf.level was around the difference 
```

```{r}
#create a new variable the computes complete therapy as nominal variable
df <- df %>% 
  rowwise() %>%
  mutate(completed_therapy = sum(!is.na(date_followup_1), 
                                 !is.na(date_followup_2),
                                 !is.na(date_followup_3),
                                 !is.na(date_followup_4))) #>=3)

table(df$completed_therapy, df$status)

```
Using the criteria that 3 sessions or more is completed therapy, we have three that fall under loss to follow up even though they had 4 appointments.
At the moment going to lump those into completed therapy group. 

```{r}
#logistic regresion 
##age_model <- glm(status ~ age, data = df, family = 'binomial')
##summary(age_model)

#create a new variable that computes complete therapy
df <- df %>% 
  rowwise() %>%
  mutate(completed_therapy = sum(!is.na(date_followup_1), 
                                 !is.na(date_followup_2),
                                 !is.na(date_followup_3),
                                 !is.na(date_followup_4))>=3) %>%
  mutate(lost_to_followup = !completed_therapy) %>%
  mutate(completed_therapy = factor(completed_therapy, 
                                    levels = c(TRUE, FALSE), 
                                    labels = c('Completed', 'Not Completed'))) %>%
  mutate(lost_to_followup = factor(lost_to_followup, 
                                   levels = c(TRUE, FALSE),
                                   labels = c('Lost', 'Not Lost')))

table(df$completed_therapy, df$status)

#table1(~factor(gender) + age + factor(marital_status) | completed_therapy, data=df)
#table1(~factor(gender) + age + factor(marital_status) | lost_to_followup, data=df)

```
```{r}
table(df$marital_status, df$completed_therapy)
table(df$completed_therapy)

```


```{r}
#logistic regression using education as a predictor
table(df$completed_therapy)

edu_table <- table(df$edu_cat, df$completed_therapy)
chisq.test(edu_table)
```
```{r}
#logistic regression using age and marital status as predictors
age_mar_model <- glm(completed_therapy ~ age + marital_status, data = df, family = 'binomial')
summary(age_mar_model)

df <- df %>% 
  mutate(completed_therapy_logical = completed_therapy == "Completed")

age_mar_model <- glm(completed_therapy_logical ~ age + marital_status, data = df, family = 'binomial')
age_mar_model_tbl <- age_mar_model %>% 
    tbl_regression(
    #label = var_labels,
    exponentiate = TRUE
  ) |> 
  add_global_p(keep = TRUE)

age_mar_model_tbl
# What should we do about the Islamic Education piece in Education variable?
# Should I go back to the dataset and try to fix the errors?

```

```{r}
gender_model <- glm(completed_therapy ~ gender, data = df, family = 'binomial')
summary(gender_model)

gender_table <- table(df$gender, df$completed_therapy)


chisq.test(gender_table)

```

```{r}
#univariable analysis
m_education <- glm(completed_therapy ~ edu_cat, family = binomial, data = df)
summary(m_education)
tidy(m_education, exponentiate = TRUE, conf.int = TRUE)
Anova(m_education)


m_marcat <- glm(completed_therapy ~ mari_cat, family = binomial, data = df)
summary(m_marcat)
tidy(m_marcat, exponentiate = TRUE, conf.int = TRUE)
Anova(m_maritalstatus)

m_gender <- glm(completed_therapy ~ gender, family = binomial, data = df)
summary(m_gender)
tidy(m_gender, exponentiate = TRUE, conf.int = TRUE)
Anova(m_gender)

m_age <- glm(completed_therapy ~ age, family = binomial, data = df)
summary(m_age)
tidy(m_age, exponentiate = TRUE, conf.int = TRUE)
Anova(m_age)

```
```{r}
#multivariable analysis
#completed_therapy, age, gender, marital status and education

mv1 <- glm(completed_therapy ~ age + gender + edu_cat + mari_cat,
           family = binomial,
           data = df)

# Extract estimates
tidy(mv1, exponentiate = TRUE, conf.int = TRUE)
# Obtain global p value
Anova(mv1)

# Create publication-ready table 
# Create a named vector for labels
var_labels <-  c(
  age = "Age in Years",
  gender = "Gender",
  edu_cat = "Highest Education Level",
  mari_cat = "Marital Status",
  phq4_score = "Diagnosis at Mental Health Assessment",
  completed_therapy = "Therapy Status of the Patient"
)

mv <- mv1 |> 
  tbl_regression(
    label = var_labels,
    exponentiate = TRUE
  ) |> 
  add_global_p()

mv
```



```{r}

```

Use the code from the IRD workshop to create tables which means I don't need to apply the log_OR and log CI codes since they are included in that but it is helpful to know how that works. 

Create a binomial variable for completed therapy

